---
title: "Homework 2"
author: "Pablo Bello"
date: '2023-01-20'
output: html_document
---

```{r setup, include=FALSE}
knitr::opts_chunk$set(echo = TRUE, message = FALSE, warning = FALSE)
```

For this assignment you are going to use the bikes dataset we began analyzing in class. You can get that dataset from the bayesrules package. Once bayesrules is installed, you can begin.

```{r}
library(tidyverse)
library(moderndive)
library(corrr)
theme_set(theme_minimal())

```


```{r}
data(bikes, package = "bayesrules")
glimpse(bikes)

```


#### Question 0

Before analyzing a dataset, it's good to know some stuff about it. Answer the following questions:

how many rows are in the dataset? 
what does each row represent?
what dates does the dataset cover?
what is the highest observed ridership in the dataset?
what was the highest wind speed recorded in the dataset?


```{r}
nrow(bikes)

tbl <- bikes

dates <- 
  tbl |> 
  summarise(min(date), max(date))

rides <- 
  tbl |> 
  slice_max(rides)

rides <- 
  tbl |> 
  slice_max(windspeed)
```

**The data has 500 rows with information about use iformation of the Capital Bikeshare service in Washington, D.C.. Contains data from 2011-01-01 to 2012-12-31.The maximum number of rides is 6946, on 2012-09-26 and a maximum windspeed recorded of 34 mph on 2011-02-19.**


### Question 1

What is the correlation between number of rides and what the temperature feels like (in Fahrenheit)? What is the correlation between the number of rides and wind speed (miles per hour)?

```{r}
tbl |> 
  select(temp_feel,rides) |> 
  correlate()

tbl |> 
  select(windspeed,rides) |> 
  correlate()
```

### Question 2

Using the approximation that a mile is equal to 1.61 kilometers, convert windspeed to kilometers per hour. Call the new variable wind_kph and add it to the bikes data frame. What is the correlation between wind speed in MPH and wind speed in KPH? Explain why in enough detail that I know you understand.

```{r}
tbl <- 
  tbl |> 
  mutate(wind_kph = 1.61*windspeed) 

tbl |> 
  select(wind_kph, windspeed) |> 
  correlate()
```

**It is the same measure scaled by a constant factor (1.61).**


### Question 3

Estimate two simple regressions:

predict rides as a function of wind speed in MPH
predict rides as a function of wind speed in KPH

Use get_regression_table() or broom::tidy() to display the results. This will give you two sets of an intercept and a slope. If any coefficients are the same between models, explain why. If any coefficients are different between models, explain why. Make sure to give me enough detail to convince me you understand.


```{r}
m1 <- lm(rides ~ windspeed,data = tbl)
m2 <- lm(rides ~ wind_kph,data = tbl)

broom::tidy(m1)
broom::tidy(m2)

```

**The intercepts are the same because scaling predictors does not change the expected value for the outcome when the predictor is set to 0. The coefficients for windspeed are scaled versions of each other. The change in rides predicted by the model with a change in the speed of the wind of 1km/hour can be transformed into the coefficient for miles per hour if multiplied by 1.61.**


### Question 4

Using the models from above, tell me what the predicted ridership would be if the wind is blowing at 20 KPH. What would the predicted ridership be if the wind is blowing at 20 MPH?

**For km/hour would be 4205 + -34.5*20 = 3515 rides. For m/hour it would be 4205 + -55.5*20 = 3095 rides.**

### Question 5

Let's get more realistic and move to multiple regression. We're going to use temp_feel in addition to wind speed. But we don't want to use Fahrenheit. So make a new variable called temp_c that is a conversion of temp_feel to Celsius and add it to the bikes dataframe. (You may have to look up how to do this conversion.)

```{r}
tbl <- 
  tbl |> 
  mutate(temp_c = (temp_feel - 32) * 5/9)


```

With this new variable, estimate a multiple regression that predicts rides as an additive function of wind_kph and temp_c (that is, no interaction). Interpret both of the slope coefficients using the following language:

"For a _____ increase in _____, the model expects a _____ increase in _____."

Now interpret the intercept.


```{r}
m2 <- lm(rides ~ wind_kph + temp_c, data = tbl)
broom::tidy(m2)
```


**For a 1km/hour increase in windspeed the model predicts on average a decrease of 19.8 rides. For a 1-degree celsius increase in temperature, the model predicts on average an increase of 143 rides. For a temperature of 0 degrees celsius and no wind, the model predicts 942 daily rides.**



### Question 6 

Using the multiple regression you just estimated, give me ridership predictions for the following situations:

SITUATION 1: temp = 25C, wind = 15 KPH
SITUATION 2: temp = 15C, wind = 5 KPH
SITUATION 3: temp = 10C, wind = 40 KPH
You can do this manually (using R as a calculator) or you might find the following code useful:

```{r}
pred_df <- tibble(
  situation = 1:3,
  temp_c = c(25, 15, 10),
  wind_kph = c(15, 5, 40),
  int = 942,
  wind_kph_beta = -19.8,
  temp_c_beta = 143
)

pred_df |> 
  mutate(pred_rides = int + temp_c_beta*temp_c + wind_kph_beta*wind_kph)
```


### Question 7

Let's add another predictor into the mix. Estimate a new model that uses weekend in addition to the predictors already in the model. Display the model results. Interpret the coefficient on weekend using a complete sentence.

```{r}
m2 <- lm(rides ~ wind_kph + temp_c + weekend, data = tbl)
broom::tidy(m2)
```

**The model predicts less rides on the weekends, with an average conditional difference of 714 rides.**

### Question 8 

If the temperature and the wind speed are average, what is the expected ridership for a weekend day? What is the expected ridership for a weekday? Show the code that gets you your answers.

```{r}
tbl <- 
  tbl |> 
  as_tibble() |> 
  mutate(wind_kph_c = wind_kph - mean(wind_kph),
         temp_c_c = temp_c - mean(temp_c)) 
  
m3 <- lm(rides ~ wind_kph_c + temp_c_c + weekend, data = tbl)
broom::tidy(m3)

```


**For an average windspeed and average temperature the expected number of rides is 3683 on weekdays and 3683 - 714 = 2969 on the weekend.**

### Question 9 

You can use get_regression_points() or predict() to see how the model did at predicting each individual value of rides. Use one of these functions to find the date with the largest absolute residual. That is, find the day where the model is most wrong about predicted ridership. Why is the model so wrong about this day? (There is a correct answer here.)

```{r}
tbl |> 
  mutate(preds = predict(m3)) |> 
  mutate(res = abs(preds - rides)) |> 
  slice_max(res) |> 
  select(date, rides, preds, res)

```

**That is the day Hurrican Sandy hit Washington D.C. and still some 20 people rode their bikes that day.**
